Planck and the Extra-terrestrials\\ or what the new SI means for astronomy

### 2020-04-10 spg/main.tex
adding an example

In the 120 years since Planck speculated about it, we have yet to make
contact with non-human or extra-terrestrial cultures, and ask them if
they indeed share Planck's views on units.  What {\em has} happened,
however, is that physical constants have turned out to have meanings
not even Planck in 1899 had imagined.  As we all know, the following
year Planck himself discovered that $h$ (formerly known as $b$) was
about quantization.  A few years later the formal relation between
mass and energy through $c^2$ turned out to be real.  Starting in the
1920s $(\hbox{Planck mass})^{-2}\times(\hbox{proton mass})^3$ was
discovered to be the mass scale of stars.  Later in the 20th century,
$2e/h$ and $h/e^2$ turned out to be macroscopically observable as
Josephson's constant and von~Klitzing's constant.  And just in the
last few years, $c^5/G$ has been observed to be the power-output scale
for merging black holes.  We can only wonder what more surprises await
us.

@@

\bibliographystyle{apsrev4-1}
\bibliography{refs.bib}

>>>>

To illustrate working with the new SI in astronomy, let us consider an
example that sounds like science fiction, but is in fact a serious
design study: the solar gravity lens
mission.\cite{2020arXiv200211871T} It envisions a telescope larger
than the solar system with the sun's gravitational field as its main
lens, for mapping the surfaces of extra-solar planets.  The essential
physical effect is that light rays going past the sun with impact
parameter $R$ are gravitationally deflected by an angle
$4GM_\odot/(c^2 R)$.  In particular, the deflection of light from a
star that is barely covered by the sun will make that star appear to
be just outside the rim of the sun.  The deflection at the rim of the
sun is well known to be $1.75\unit{arcsec}$ and its first measurement
a century ago is the stuff of legend.\cite{Kennefick2012} Knowing the
solar mass in gravity-seconds and approximating the solar radius as
$R_\odot\approx c\times2\unit{s}$, we can trivially work out the
gravitational deflection angle $\alpha=4GM_\odot/(c^2 R_\odot)$ as
$\approx10^{-5}\unit{radians}$, which is indeed a little under
$2\unit{arcsec}$.  If we go far-enough from the sun that its apparent
radius on the sky becomes smaller than the deflection angle, a
remarkable phenomenon will be seen: light sources precisely behind the
sun will appear as luminous rings (known as Einstein rings) around the
sun.  The threshold distance to see an Einstein ring is
$R_\odot/\alpha\approx c\times2\times10^5\unit{s}$ or $\approx
400\unit{au}$.  (Actually about $550\unit{au}$ with less drastic
approximations.)

$g_{\odot/\oplus} \approx 6\times10^{-3}\unit{m}\unit{s}^{-2}$

$P_{\odot/\oplus} \approx 10^{-5}\unit{Pa}$

Would need $\approx 600\unit{m}^2\unit{kg}^{-1}$ to make the sun a
net repeller.

Cunning hybrid.

In the 120 years since Planck speculated about it, we have yet to make
contact with non-human or extra-terrestrial cultures, and ask them if
they indeed share Planck's views on units.  What {\em has} happened,
however, is that physical constants have turned out to have meanings
not even Planck in 1899 had imagined.  As we all know, the following
year Planck himself discovered that $h$ (formerly known as $b$) was
about quantization.  A few years later the formal relation between
mass and energy through $c^2$ turned out to be real.  Starting in the
1920s $(hc/G)^{-2}\times(\hbox{proton mass})^3$ was
discovered to be the mass scale of stars.  Later in the 20th century,
$2e/h$ and $h/e^2$ turned out to be macroscopically observable as
Josephson's constant and von~Klitzing's constant.  And just in the
last few years, $c^5/G$ has been observed to be the power-output scale
for merging black holes.  We can only wonder what more surprises await
us.

@@

\bibliographystyle{apsrev4-1}
%\bibliography{main.bbl}
\bibliography{refs.bib}
